---
title: "Robustness analysis"
output: html_document
author: "Matthew Versace, Sharna McClure & Ciara Stevens-Baytopp"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE
)

library(kableExtra)
library(tidyverse)
library(readrba)
library(PerformanceAnalytics)
library(quantmod)
library(lmtest)
library(sandwich)
library(zoo)

# All your data download and model code here
# Remove ALL head(), tail(), cat(), print(), summary() calls
# These are for console checking only — not needed in the Rmd

options("getSymbols.warning4.0" = FALSE)
getSymbols(Symbols = c("CBA.AX", "BHP.AX", "WES.AX", "CSL.AX"),
           from = "2014-01-01", to = "2024-12-31",
           auto.assign = TRUE, warnings = FALSE)
getSymbols(Symbols = "^AXJO", from = "2014-01-01", to = "2024-12-31",
           auto.assign = TRUE, warnings = FALSE)

cba_D_P  <- to.daily(CBA.AX$CBA.AX.Adjusted, OHLC = FALSE)
bhp_D_P  <- to.daily(BHP.AX$BHP.AX.Adjusted, OHLC = FALSE)
wes_D_P  <- to.daily(WES.AX$WES.AX.Adjusted, OHLC = FALSE)
csl_D_P  <- to.daily(CSL.AX$CSL.AX.Adjusted, OHLC = FALSE)
axjo_D_P <- to.daily(AXJO$AXJO.Adjusted,      OHLC = FALSE)
colnames(axjo_D_P) <- "ASX200"

daily_P <- merge(cba_D_P, bhp_D_P, wes_D_P, csl_D_P)
colnames(daily_P) <- c("cba", "bhp", "wes", "csl")

lr <- na.omit(diff(log(daily_P)))
colnames(lr) <- c("cba", "bhp", "wes", "csl")

market_lr <- na.omit(diff(log(axjo_D_P)))
colnames(market_lr) <- "ASX200"

common_dates <- intersect(index(lr), index(market_lr))
lr           <- lr[common_dates]
market_lr    <- market_lr[common_dates]

reg_data <- data.frame(
  date      = index(lr),
  coredata(lr),
  market_lr = coredata(market_lr)
)
colnames(reg_data) <- c("date", "cba", "bhp", "wes", "csl", "market_lr")

for (i in seq_along(colnames(lr))) {
  name <- colnames(lr)[i]
  ret  <- reg_data[[name]]
  sd_r <- sd(ret, na.rm = TRUE)
  reg_data[[paste0(name, "_momentum")]] <- lag(
    rollsum(ret, k = 20, fill = NA, align = "right"), 1)
  reg_data[[paste0(name, "_large_pos")]] <- lag(
    ifelse(ret >  2 * sd_r, 1, 0), 1)
  reg_data[[paste0(name, "_large_neg")]] <- lag(
    ifelse(ret < -2 * sd_r, 1, 0), 1)
}

rba_rate <- read_rba(series_id = "FIRMMCRTD") %>%
  select(date, value) %>%
  rename(cash_rate = value) %>%
  mutate(date = as.Date(date)) %>%
  filter(date >= as.Date("2014-01-01") &
           date <= as.Date("2024-12-31")) %>%
  arrange(date) %>%
  mutate(
    rate_change = cash_rate - lag(cash_rate),
    hike        = ifelse(rate_change > 0, 1, 0),
    cut         = ifelse(rate_change < 0, 1, 0),
    change      = ifelse(rate_change != 0, 1, 0)
  )

reg_data <- reg_data %>%
  left_join(rba_rate %>%
              select(date, rate_change, hike, cut, change),
            by = "date") %>%
  mutate(
    rate_change = replace_na(rate_change, 0),
    hike        = replace_na(hike, 0),
    cut         = replace_na(cut, 0),
    change      = replace_na(change, 0)
  ) %>%
  mutate(
    hike_post1 = lag(hike, 1, default = 0),
    hike_post2 = lag(hike, 2, default = 0),
    cut_post1  = lag(cut,  1, default = 0),
    cut_post2  = lag(cut,  2, default = 0)
  )

rf_data <- read_rba(series_id = "FIRMMBAB90") %>%
  select(date, value) %>%
  rename(rf_annual = value) %>%
  mutate(date = as.Date(date),
         rf_daily = rf_annual / 100 / 252) %>%
  filter(date >= as.Date("2014-01-01") &
           date <= as.Date("2024-12-31"))

reg_data <- reg_data %>%
  left_join(rf_data %>% select(date, rf_daily), by = "date") %>%
  fill(rf_daily, .direction = "down")

reg_data$cba_excess    <- reg_data$cba       - reg_data$rf_daily
reg_data$bhp_excess    <- reg_data$bhp       - reg_data$rf_daily
reg_data$wes_excess    <- reg_data$wes       - reg_data$rf_daily
reg_data$csl_excess    <- reg_data$csl       - reg_data$rf_daily
reg_data$market_excess <- reg_data$market_lr - reg_data$rf_daily

capm_cba <- lm(cba_excess ~ market_excess + hike + cut +
                 hike_post1 + cut_post1 + cba_momentum +
                 cba_large_pos + cba_large_neg, data = reg_data)
capm_bhp <- lm(bhp_excess ~ market_excess + hike + cut +
                 hike_post1 + cut_post1 + bhp_momentum +
                 bhp_large_pos + bhp_large_neg, data = reg_data)
capm_wes <- lm(wes_excess ~ market_excess + hike + cut +
                 hike_post1 + cut_post1 + wes_momentum +
                 wes_large_pos + wes_large_neg, data = reg_data)
capm_csl <- lm(csl_excess ~ market_excess + hike + cut +
                 hike_post1 + cut_post1 + csl_momentum +
                 csl_large_pos + csl_large_neg, data = reg_data)

T_cba  <- nobs(capm_cba)
m_cba  <- floor(0.75 * T_cba^(1/3))
NW_cba <- NeweyWest(capm_cba, lag = m_cba - 1, prewhite = FALSE, adjust = TRUE)

T_bhp  <- nobs(capm_bhp)
m_bhp  <- floor(0.75 * T_bhp^(1/3))
NW_bhp <- NeweyWest(capm_bhp, lag = m_bhp - 1, prewhite = FALSE, adjust = TRUE)

T_wes  <- nobs(capm_wes)
m_wes  <- floor(0.75 * T_wes^(1/3))
NW_wes <- NeweyWest(capm_wes, lag = m_wes - 1, prewhite = FALSE, adjust = TRUE)

T_csl  <- nobs(capm_csl)
m_csl  <- floor(0.75 * T_csl^(1/3))
NW_csl <- NeweyWest(capm_csl, lag = m_csl - 1, prewhite = FALSE, adjust = TRUE)

reg_post <- reg_data %>%
  filter(date >= as.Date("2021-01-01")) %>%
  select(-cut, -cut_post1)

capm_cba_post <- lm(cba_excess ~ market_excess + hike +
                      hike_post1 + cba_momentum +
                      cba_large_pos + cba_large_neg, data = reg_post)
capm_bhp_post <- lm(bhp_excess ~ market_excess + hike +
                      hike_post1 + bhp_momentum +
                      bhp_large_pos + bhp_large_neg, data = reg_post)
capm_wes_post <- lm(wes_excess ~ market_excess + hike +
                      hike_post1 + wes_momentum +
                      wes_large_pos + wes_large_neg, data = reg_post)
capm_csl_post <- lm(csl_excess ~ market_excess + hike +
                      hike_post1 + csl_momentum +
                      csl_large_pos + csl_large_neg, data = reg_post)

capm_cba_mag <- lm(cba_excess ~ market_excess + rate_change +
                     hike_post1 + cut_post1 + cba_momentum +
                     cba_large_pos + cba_large_neg, data = reg_data)
capm_bhp_mag <- lm(bhp_excess ~ market_excess + rate_change +
                     hike_post1 + cut_post1 + bhp_momentum +
                     bhp_large_pos + bhp_large_neg, data = reg_data)
capm_wes_mag <- lm(wes_excess ~ market_excess + rate_change +
                     hike_post1 + cut_post1 + wes_momentum +
                     wes_large_pos + wes_large_neg, data = reg_data)
capm_csl_mag <- lm(csl_excess ~ market_excess + rate_change +
                     hike_post1 + cut_post1 + csl_momentum +
                     csl_large_pos + csl_large_neg, data = reg_data)

format_coef <- function(model, row) {
  coefs <- summary(model)$coefficients
  if (row > nrow(coefs)) return("—")
  est   <- formatC(coefs[row, 1], format = "f", digits = 6)
  se    <- formatC(coefs[row, 2], format = "f", digits = 4)
  p     <- coefs[row, 4]
  stars <- ifelse(p < 0.001, "***",
           ifelse(p < 0.01,  "**",
           ifelse(p < 0.05,  "*",
           ifelse(p < 0.10,  ".", ""))))
  paste0(est, stars, " (", se, ")")
}

format_hac <- function(model, vcov_mat, row) {
  coefs <- coeftest(model, vcov = vcov_mat)
  if (row > nrow(coefs)) return("—")
  est   <- formatC(coefs[row, 1], format = "f", digits = 6)
  se    <- formatC(coefs[row, 2], format = "f", digits = 4)
  p     <- coefs[row, 4]
  stars <- ifelse(p < 0.001, "***",
           ifelse(p < 0.01,  "**",
           ifelse(p < 0.05,  "*",
           ifelse(p < 0.10,  ".", ""))))
  paste0(est, stars, " (", se, ")")
}

format_coef_post <- function(model, row) {
  coefs <- summary(model)$coefficients
  if (row > nrow(coefs)) return("—")
  est   <- formatC(coefs[row, 1], format = "f", digits = 6)
  se    <- formatC(coefs[row, 2], format = "f", digits = 4)
  p     <- coefs[row, 4]
  stars <- ifelse(p < 0.001, "***",
           ifelse(p < 0.01,  "**",
           ifelse(p < 0.05,  "*",
           ifelse(p < 0.10,  ".", ""))))
  paste0(est, stars, " (", se, ")")
}

```

## Analysis Results 

This descriptive analysis estimated the CAPM model for four ASX-listed stocks over the ten year period of 2014-2024, testing the Efficient Market Hypothesis (EMH) through RBA dummy variables, and behavioural variables. Some of the key findings include a significant alpha value for WES, implying mispricing of the asset relative to the level of systematic risk, delayed reaction to prices for CBA after a rate hike, overreaction in CSL, and regression to the mean in momentum for BHP. All of these results are estimates of correlations and not claims of causality. These correlations show consistency with semi-strong form of efficient market hypothesis for all stocks, and the effect of new information from RBA produces heterogeneous return effects across stocks, with some exhibiting greater sensitivity to interest rate changes compared to others.

## Robustness Analysis  

### Test 1: Newey-West

The Newey-West test is a correction to autocorrelation and heteroskedasticity by providing a consistent covariance estimator. Coefficients will remain the same, but standard errors, t-values, and p-values are computed again.

While it is commonly assumed that financial data has a moderate risk to autocorrelation and a substantial risk to heteroskedasticity, we formally test for these by employing a Breusch-Godfrey and a White auxiliary regression model (white test) for all stocks.

The White and Breusch-Godfrey tests produced the following results:

Figure 1: White and Breusch-Godfrey tests (Beginning with White test)

| Stock      | F-statistic                   | p-value      |
|:-----------|:------------------------------|:-------------|
| CBA        | 18.06                         | \< 0.001     |
| BHP        | 34.95                         | \< 0.001     |
| WES        | 11.97                         | \< 0.001     |
| CSL        | 16.16                         | \< 0.001     |
| **Stock**  | **Lagrange Multiplier value** | **p-value**  |
| CBA        | 8.823                         | 0.0012       |
| BHP        | 9.762                         | 0.008        |
| WES        | 4.592                         | 0.101        |
| CSL        | 0.378                         | 0.828        |

```{r results table, echo=FALSE}
var_names_full <- c("Alpha (Intercept)",
                    "Market Excess Return (β)",
                    "Rate Hike",
                    "Rate Cut",
                    "Post Hike Day 1",
                    "Post Cut Day 1",
                    "Momentum",
                    "Overreaction (+)",
                    "Overreaction (-)")

rob_hac_table <- data.frame(
  Variable  = var_names_full,
  CBA_OLS   = sapply(1:9, function(i) format_coef(capm_cba, i)),
  CBA_HAC   = sapply(1:9, function(i) format_hac(capm_cba, NW_cba, i)),
  BHP_OLS   = sapply(1:9, function(i) format_coef(capm_bhp, i)),
  BHP_HAC   = sapply(1:9, function(i) format_hac(capm_bhp, NW_bhp, i)),
  WES_OLS   = sapply(1:9, function(i) format_coef(capm_wes, i)),
  WES_HAC   = sapply(1:9, function(i) format_hac(capm_wes, NW_wes, i)),
  CSL_OLS   = sapply(1:9, function(i) format_coef(capm_csl, i)),
  CSL_HAC   = sapply(1:9, function(i) format_hac(capm_csl, NW_csl, i)),
  stringsAsFactors = FALSE
)

r2_row_hac <- data.frame(
  Variable  = "R²",
  CBA_OLS   = as.character(round(summary(capm_cba)$r.squared, 3)),
  CBA_HAC   = as.character(round(summary(capm_cba)$r.squared, 3)),
  BHP_OLS   = as.character(round(summary(capm_bhp)$r.squared, 3)),
  BHP_HAC   = as.character(round(summary(capm_bhp)$r.squared, 3)),
  WES_OLS   = as.character(round(summary(capm_wes)$r.squared, 3)),
  WES_HAC   = as.character(round(summary(capm_wes)$r.squared, 3)),
  CSL_OLS   = as.character(round(summary(capm_csl)$r.squared, 3)),
  CSL_HAC   = as.character(round(summary(capm_csl)$r.squared, 3)),
  stringsAsFactors = FALSE
)

obs_row_hac <- data.frame(
  Variable  = "Observations",
  CBA_OLS   = as.character(nobs(capm_cba)),
  CBA_HAC   = as.character(nobs(capm_cba)),
  BHP_OLS   = as.character(nobs(capm_bhp)),
  BHP_HAC   = as.character(nobs(capm_bhp)),
  WES_OLS   = as.character(nobs(capm_wes)),
  WES_HAC   = as.character(nobs(capm_wes)),
  CSL_OLS   = as.character(nobs(capm_csl)),
  CSL_HAC   = as.character(nobs(capm_csl)),
  stringsAsFactors = FALSE
)

final_rob_hac <- bind_rows(rob_hac_table, r2_row_hac, obs_row_hac)

final_rob_hac %>%
  kable(
    caption   = "Figure 2: Robustness Check — Newey-West HAC Standard Errors",
    align     = c("l", rep("c", 8)),
    col.names = c("Variable",
                  "CBA (OLS)", "CBA (HAC)",
                  "BHP (OLS)", "BHP (HAC)",
                  "WES (OLS)", "WES (HAC)",
                  "CSL (OLS)", "CSL (HAC)")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
    full_width        = TRUE,
    font_size         = 11
  ) %>%
  row_spec(0, bold = TRUE, background = "#2c7bb6", color = "white") %>%
  row_spec(nrow(final_rob_hac) - 1,
           extra_css = "border-top: 2px solid black;") %>%
  row_spec(nrow(final_rob_hac),
           extra_css = "border-bottom: 2px solid black;") %>%
  column_spec(1, bold = TRUE, width = "8em") %>%
  add_header_above(c(" " = 1,
                     "CBA" = 2,
                     "BHP" = 2,
                     "WES" = 2,
                     "CSL" = 2)) %>%
  scroll_box(width = "100%") %>%
  add_footnote(
    c("Column 1 (OLS) is the main specification using standard OLS standard errors.",
      "Column 2 (HAC) applies Newey-West correction.",
      "R² identical across OLS and HAC — HAC only affects standard errors not point estimates.",
      "Standard errors in parentheses.",
      "*** p<0.001, ** p<0.01, * p<0.05"),
    notation = "none"
  )

```

From Figure 1, we reject the null hypothesis of homoskedasticity for all four stocks and confirm the original observation. Clusters of volatility are seen throughout the model which could explain why the regression residuals are heteroskedastic. The Breusch-Godfrey test revealed CBA and BHP to have autocorrelation with a significant p-value, while WES and CSL do not present this.

### Result

Finally, applying the Newey-West to the CAPM models altered the significance of several findings. All beta values, and the WES alpha at the 5% level (p = 0.011) remain significant, representing the most robust finding. However, every other finding in the regression analysis lost its statistical significance. Significant beta values confirm that systematic risk has been robustly priced into the model. CBA's delayed rate hike reaction and BHP's regression to the mean are likely sensitive to autocorrelation following the Breusch-Godfrey results. The most interesting finding was CSL's reaction variable; while the p-values show no significance, the coefficient signs remain the same. This implies that the original CAPM for CSL overestimated the magnitude of the overreaction to changes in the stock prices since the direction of the effect appears to be consistent. Despite losing many significant values, this result suggests the original CAPM results have been overstated, and conclusions regarding EMH should be considered with caution. The robust WES finding and consistent CSL direction still imply semi strong-form EMH for these stocks. However, CBA and BHP show no significant behavioural or RBA coefficients, providing no evidence against EMH.

### Test 2: Post-COVID Sample Period

```{r, echo=FALSE}
format_coef <- function(model, row) {
  coefs <- summary(model)$coefficients
  if (row > nrow(coefs)) return("—")
  est   <- formatC(coefs[row, 1], format = "f", digits = 6)
  se    <- formatC(coefs[row, 2], format = "f", digits = 4)
  p     <- coefs[row, 4]
  stars <- ifelse(p < 0.001, "***",
           ifelse(p < 0.01,  "**",
           ifelse(p < 0.05,  "*",
           ifelse(p < 0.10,  ".", ""))))
  p_fmt <- ifelse(p < 0.001, "<0.001",
                  formatC(p, format = "f", digits = 3))
  paste0(est, stars, " (", se, ")")
}

rob_post_table <- data.frame(
  Variable   = var_names_full,
  CBA_Full   = sapply(1:9, function(i) format_coef(capm_cba,      i)),
  CBA_Post   = c(
    format_coef_post(capm_cba_post, 1),
    format_coef_post(capm_cba_post, 2),
    format_coef_post(capm_cba_post, 3),
    "— (no cuts)",
    format_coef_post(capm_cba_post, 4),
    "— (no cuts)",
    format_coef_post(capm_cba_post, 5),
    format_coef_post(capm_cba_post, 6),
    format_coef_post(capm_cba_post, 7)
  ),
  BHP_Full   = sapply(1:9, function(i) format_coef(capm_bhp,      i)),
  BHP_Post   = c(
    format_coef_post(capm_bhp_post, 1),
    format_coef_post(capm_bhp_post, 2),
    format_coef_post(capm_bhp_post, 3),
    "— (no cuts)",
    format_coef_post(capm_bhp_post, 4),
    "— (no cuts)",
    format_coef_post(capm_bhp_post, 5),
    format_coef_post(capm_bhp_post, 6),
    format_coef_post(capm_bhp_post, 7)
  ),
  WES_Full   = sapply(1:9, function(i) format_coef(capm_wes,      i)),
  WES_Post   = c(
    format_coef_post(capm_wes_post, 1),
    format_coef_post(capm_wes_post, 2),
    format_coef_post(capm_wes_post, 3),
    "— (no cuts)",
    format_coef_post(capm_wes_post, 4),
    "— (no cuts)",
    format_coef_post(capm_wes_post, 5),
    format_coef_post(capm_wes_post, 6),
    format_coef_post(capm_wes_post, 7)
  ),
  CSL_Full   = sapply(1:9, function(i) format_coef(capm_csl,      i)),
  CSL_Post   = c(
    format_coef_post(capm_csl_post, 1),
    format_coef_post(capm_csl_post, 2),
    format_coef_post(capm_csl_post, 3),
    "— (no cuts)",
    format_coef_post(capm_csl_post, 4),
    "— (no cuts)",
    format_coef_post(capm_csl_post, 5),
    format_coef_post(capm_csl_post, 6),
    format_coef_post(capm_csl_post, 7)
  ),
  stringsAsFactors = FALSE
)

r2_row_post <- data.frame(
  Variable  = "R²",
  CBA_Full  = as.character(round(summary(capm_cba)$r.squared,      3)),
  CBA_Post  = as.character(round(summary(capm_cba_post)$r.squared,  3)),
  BHP_Full  = as.character(round(summary(capm_bhp)$r.squared,      3)),
  BHP_Post  = as.character(round(summary(capm_bhp_post)$r.squared,  3)),
  WES_Full  = as.character(round(summary(capm_wes)$r.squared,      3)),
  WES_Post  = as.character(round(summary(capm_wes_post)$r.squared,  3)),
  CSL_Full  = as.character(round(summary(capm_csl)$r.squared,      3)),
  CSL_Post  = as.character(round(summary(capm_csl_post)$r.squared,  3)),
  stringsAsFactors = FALSE
)

obs_row_post <- data.frame(
  Variable  = "Observations",
  CBA_Full  = as.character(nobs(capm_cba)),
  CBA_Post  = as.character(nobs(capm_cba_post)),
  BHP_Full  = as.character(nobs(capm_bhp)),
  BHP_Post  = as.character(nobs(capm_bhp_post)),
  WES_Full  = as.character(nobs(capm_wes)),
  WES_Post  = as.character(nobs(capm_wes_post)),
  CSL_Full  = as.character(nobs(capm_csl)),
  CSL_Post  = as.character(nobs(capm_csl_post)),
  stringsAsFactors = FALSE
)

final_rob_post <- bind_rows(rob_post_table, r2_row_post, obs_row_post)

final_rob_post %>%
  kable(
    caption   = "Figure 3: Robustness Check — Post-COVID Subsample (2021-2024)",
    align     = c("l", rep("c", 8)),
    col.names = c("Variable",
                  "CBA (Full)", "CBA (Post)",
                  "BHP (Full)", "BHP (Post)",
                  "WES (Full)", "WES (Post)",
                  "CSL (Full)", "CSL (Post)")
  ) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "bordered"),
    full_width        = TRUE,
    font_size         = 11
  ) %>%
  row_spec(0, bold = TRUE, background = "#2c7bb6", color = "white") %>%
  row_spec(nrow(final_rob_post) - 1,
           extra_css = "border-top: 2px solid black;") %>%
  row_spec(nrow(final_rob_post),
           extra_css = "border-bottom: 2px solid black;") %>%
  column_spec(1, bold = TRUE, width = "8em") %>%
  add_header_above(c(" " = 1,
                     "CBA" = 2,
                     "BHP" = 2,
                     "WES" = 2,
                     "CSL" = 2)) %>%
  scroll_box(width = "100%") %>%
  add_footnote(
    c("Column 1 (Full) is the main specification using full sample 2014-2024.",
      "Column 2 (Post) restricts sample to January 2021 - December 2024.",
      "Cut variables excluded from post-COVID — no RBA cuts occurred in this period.",
      "Standard errors in parentheses.",
      "*** p<0.001, ** p<0.01, * p<0.05"),
    notation = "none"
  )

```

We estimate the CAPM model from the beginning of the RBA rate rise cycle (Post-COVID period, 01/01/2021) to see the impact it has on the stocks. These results reveal heterogeneity in consideration of the original CAPM model. CBA demonstrates consistency when compared to the original model, with delayed price adjustment following the rate rises as previously reported. Interestingly, the CBA alpha value is now positive at the 1% level (p=0.00677), further reflecting a semi-strong form of market inefficiency under the post-COVID period. In contrast to the original model, WES and CSL lose the statistical significance previously observed. This illustrates that CSL's overreaction was likely driven by broader volatility of the full sample period, and WES's alpha was concentrated in a pre-COVID sample period. BHP now exhibits a significant large negative behavioural observation. Days that were substantially negative during the post-COVID instability were followed by more negative returns. Concluding, these results confirm that some conditional correlations are driven by sub-periods. This raises slight concern for the original CAPM's validity as it may overstate market inefficiency because of the restriction in sample as this change reveals a more efficient market for CSL and WES.

### Test 3: Alternative Specification 

Figure 4 (Alternative Specification Results)

| Stocks  | Rate Change Coefficients  | p-value  |
|---------|---------------------------|----------|
| CBA     | -0.0071628                | 0.1906   |
| BHP     | 0.0014724                 | 0.8756   |
| WES     | -0.0004656                | 0.9430   |
| CSL     | 0.0066622                 | 0.41745  |

One of the threats to our analysis was the use of a dummy information variable for RBA rate cuts and hikes, where magnitude was not considered. In this test we examine whether the magnitude of the hike (change in basis points) has any impact on results. To do this, we estimate the CAPM again, including the new rate_change variable which includes actual RBA interest rate values. From the results, none of the new coefficient p-values show significance. This could suggest that the market only reacts to the decision itself (raise/cut) rather than the size of the change in basis points. Another finding we discovered was a marginally significant alpha value for CSL at the 10% level of significance. This finding does not meet the conventional 5% level used in this analysis and should be interpreted with caution. Values of significance found in the original CAPM still hold true and thus survived this test, with minor or no change to model fitness in R-squared.

## Conclusion

This robustness analysis stress tested the CAPM model using alternative sample and inference tests. We employed the Newey-West HAC, estimated the CAPM during the post-COVID rate hike period where the RBA increased interest rates 13 times, and measured the magnitude of interest rate changes. As anticipated, beta remained consistently significant across the three checks. The most robust findings of this analysis appear to be the CBA post-hike delayed price adjustment and the WES alpha, with both results surviving one of three tests and providing partially robust evidence of a market inefficiency. The WES alpha survived the Newey-West HAC, but failed in the post-COVID sub-sample, implying that different dynamics including (the RBA's historically low interest rate environment) could have influenced results. CBA's post hike did not survive HAC, but was more successful in the alternate sample, suggesting RBA rate rise information has a larger impact. Behavioural findings of CSL's overreaction and BHP's regression to the mean (momentum) appear to be sensitive to both checks, indicating questionable validity for the analysis. The heterogeneity of these results provides a useful conclusion that can be made about the broader concept of EMH and the impact of new information on stocks. Consistent with financial theory, EMH and new information will naturally not apply to all securities, with some following random walk, and others being predictable. Additionally, this conclusion is amplified by behavioural biases that influence investor behaviour for some stocks, influencing return variability. It can therefore be claimed that while conditional correlations with new information and stock returns exist, they only apply under specific circumstances and are sensitive to inference assumptions. This is broadly consistent with a market that is mostly but not perfect efficient.
